---
title: "Image Classification with R"
output: html_notebook
---
The objective of this project is to classify large amounts of images using R. The classification is to be made using 4 different models:

* Naïve Bayes
* Random Forest
* Deep Learning
* Support Vector Machines

Load Required Libraries
```{r}
library(imager)
library(tidyverse)
library(h2o)
library(kernlab)
```


#### Formula

Formula to generate a single row of 1 label and 75 features for an image
```{r}
getNumbers <- function(image,label){
# Load image 
im <- load.image(image)
# Detect edges and set standard size
im <- deriche(im,2,order=2,axis="x") %>% deriche(2,order=2,axis="y") %>% resize(50,50)
# Convert to Data Frame
df <- as.data.frame(im)
# Add Label
row <- data.frame(Label = label)
# Separate Colors
R <- subset(df,df$cc == 1)
G <- subset(df,df$cc == 2)
B <- subset(df,df$cc == 3)
# Create Grouping Vectors
c10 <- c(1:10)
c20 <- c(11:20)
c30 <- c(21:30)
c40 <- c(31:40)
c50 <- c(41:50)

# 3 Loops, one for each Color  of RGB

# Loop through X
for(i in 1:5) {
  x <- i
  # Change the Grouping Vector depending on x itteration
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  # Loop through y
  for (j in 1:5) {
    y <- j
    # Change the Grouping Vector depending on y itteration
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    # Change name of each column 
    valName <- paste0("x", i, "y", j, ".R")
    # Group based on the coordiate's groups
    temp <-
      R %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    # Add to the row data frame
    row <- cbind(row, temp)
  }
  
}
# Repeat proces for G
for(i in 1:5) {
  x <- i
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  for (j in 1:5) {
    y <- j
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    valName <- paste0("x", i, "y", j, ".G")
    temp <-
      G %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    row <- cbind(row, temp)
  }
  
}
# Repeat Process for B
for(i in 1:5) {
  x <- i
  onX <- case_when(x == 1 ~ c10,
                   x == 2 ~ c20,
                   x == 3 ~ c30,
                   x == 4 ~ c40,
                   x == 5 ~ c50)
  for (j in 1:5) {
    y <- j
    onY <- case_when(y == 1 ~ c10,
                     y == 2 ~ c20,
                     y == 3 ~ c30,
                     y == 4 ~ c40,
                     y == 5 ~ c50)
    valName <- paste0("x", i, "y", j, ".B")
    temp <-
      B %>% filter(x %in% onX &
                     y %in% onY) %>% summarize(val = mean(value))
    colnames(temp) <- valName
    row <- cbind(row, temp)
  }
  
}
# Return Final Data Set
return(row)
}

```


#### Convert to Numbers

Loop throught all the files two separate folders, use `getNumbers()` for each image and 
```{r}
# List the files of the Cars Folder
Car_files <-
  list.files(path = "C:/Users/moise/Documents/R/STU/Spring 2020/Cars")
# Create Empty Data Frame
Cars <- data.frame()

# Loop each file
for (i in 1:length(Car_files)) {
  # Get the path for each photo
  foto <- paste0("Cars/", Car_files[i])
  # Use formula to generate the numbers and assign the label "Car"
  data <- getNumbers(foto, "Car")
  # Bind Each Result together
  Cars <- rbind(Cars, data)
}

# Repeat process for the images that are not Cars
No_Cars_files <-
  list.files(path = "C:/Users/moise/Documents/R/STU/Spring 2020/Not_Cars")
Not_Cars <- data.frame()

for (i in 1:length(No_Cars_files)) {
  foto <- paste0("Not_Cars/", No_Cars_files[i])
  data <- getNumbers(foto, "Not_Car")
  Not_Cars <- rbind(Not_Cars, data)
}
# Bind Cars and Not Cars data frames.
Data <- rbind(Cars,Not_Cars)
# Change the labels to factors
Data$Label <- as.factor(Data$Label)
```

## Models

H2o Split

```{r}
# Use all CPU threads
h2o.init(nthreads = -1)

SPLIT_RATIO = .75
h2oCars <- as.h2o(Data)
carsSplit <- h2o.splitFrame(data = h2oCars, ratios = SPLIT_RATIO)
h2oTrain <- carsSplit[[1]]
h2oTest <- carsSplit[[2]]
```

### DEEP LEARNING
```{r}
h2oDL <- h2o.deeplearning(
  x = colnames(h2oTrain),
  y = c("Label"),
  training_frame = h2oTrain,
  hidden = c(16, 8),
  epochs = 100,
  seed = 12345,
  nfolds = 3
)


pred <- h2o.predict(h2oDL, h2oTest)
result_DLh2o <- data.frame(Actual = as.vector(h2oTest$Label), Prediction = as.vector(pred$predict))

# TP,TN, FP, FN
TPDl <- subset(result_DLh2o, Actual == "Car" & Prediction == "Car" )
TNDl <- subset(result_DLh2o, Actual == "No_Car" & Prediction == "No_Car")
FPDl <- subset(result_DLh2o, Actual == "No_Car" & Prediction == "Car")
FNDl <- subset(result_DLh2o, Actual == "Car" & Prediction == "No_Car")

# Accuracy
accuracyDL <- (nrow(TPDl)) / (nrow(TPDl) + nrow(FPDl))
cat("Deep Learning Accuracy:" , accuracyDL, "\n")

# Precision
precisionDL <- (nrow(TPDl)) / (nrow(h2oTest))
cat("Deep Learning Precision:" , precisionDL, "\n")

# Recall
recallDL <- (nrow(TPDl)) / (nrow(TPDl) + nrow(FNDl))
cat("Deep Learning Recall:" , recallDL, "\n")

# Performance
rfPerformance <- h2o.performance(h2oDL, h2oTest)

# Auc
h2oDlAUC <- h2o.auc(rfPerformance)
cat("Deep Learning AuC:", h2oDlAUC, "\n")
```

### RANDOM FOREST
```{r}
rfModel <-
  h2o.randomForest(y = "Label",
                   training_frame = h2oTrain,
                   validation_frame = h2oTest)

rfPred <- h2o.predict(rfModel, h2oTest)
result_RFh2o <- data.frame(Actual = as.vector(h2oTest$Label), Prediction = as.vector(rfPred$predict))

# TP,TN, FP, FN
TPRf <- subset(result_RFh2o, Actual == "Car" & Prediction == "Car" )
TNRf <- subset(result_RFh2o, Actual == "No_Car" & Prediction == "No_Car")
FPRf <- subset(result_RFh2o, Actual == "No_Car" & Prediction == "Car")
FNRf <- subset(result_RFh2o, Actual == "Car" & Prediction == "No_Car")

# Accuracy
accuracyRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FPRf))
cat("Random Forest Accuracy:" , accuracyRf, "\n")

# Precision
precisionRf <- (nrow(TPRf)) / (nrow(h2oTest))
cat("Random Forest Precision:" , precisionRf, "\n")

# Recall
recallRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FNRf))
cat("Random Forest Recall:" , recallRf, "\n")

# Performance
rfPerformance <- h2o.performance(rfModel, h2oTest)

# Auc
h2oRfAUC <- h2o.auc(rfPerformance)
cat("Random Forest AuC:", h2oRfAUC, "\n")
```
### NAIVE BAYES
```{r}
nbModel <-
  h2o.naiveBayes(y = "Label",
                   training_frame = h2oTrain,
                   validation_frame = h2oTest)

nbPred <- h2o.predict(nbModel, h2oTest)
result_RFh2o <- data.frame(Actual = as.vector(h2oTest$Label), Prediction = as.vector(nbPred$predict))

# TP,TN, FP, FN
TPRf <- subset(result_RFh2o, Actual == "Car" & Prediction == "Car" )
TNRf <- subset(result_RFh2o, Actual == "No_Car" & Prediction == "No_Car")
FPRf <- subset(result_RFh2o, Actual == "No_Car" & Prediction == "Car")
FNRf <- subset(result_RFh2o, Actual == "Car" & Prediction == "No_Car")

# Accuracy
accuracyRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FPRf))
cat("Naive Bayes Accuracy:" , accuracyRf, "\n")

# Precision
precisionRf <- (nrow(TPRf)) / (nrow(h2oTest))
cat("Naive Bayes Precision:" , precisionRf, "\n")

# Recall
recallRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FNRf))
cat("Naive Bayes Recall:" , recallRf, "\n")

# Performance
rfPerformance <- h2o.performance(nbModel, h2oTest)

# Auc
h2oRfAUC <- h2o.auc(rfPerformance)
cat("Naive Bayes AuC:", h2oRfAUC, "\n")
```
### SVM

Regular Split
```{r}
sample <- sample.split(Data$Label, SplitRatio = .75)
train <- subset(Data, sample == TRUE)
test <- subset(Data, sample == FALSE)
```

```{r}
modelSVM <- ksvm(Label ~ ., data = train, kernel = "rbfdot")

predSVM <- predict(modelSVM,test)
result_SVM <- data.frame(Actual = test$Label, Prediction = predSVM)

# TP,TN, FP, FN
TPRf <- subset(result_SVM, Actual == "Car" & Prediction == "Car" )
TNRf <- subset(result_SVM, Actual == "No_Car" & Prediction == "No_Car")
FPRf <- subset(result_SVM, Actual == "No_Car" & Prediction == "Car")
FNRf <- subset(result_SVM, Actual == "Car" & Prediction == "No_Car")

# Accuracy
accuracyRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FPRf))
cat("Gaussian RBF Kernel Accuracy:" , accuracyRf, "\n")

# Precision
precisionRf <- (nrow(TPRf)) / (nrow(h2oTest))
cat("Gaussian RBF Kernel Precision:" , precisionRf, "\n")

# Recall
recallRf <- (nrow(TPRf)) / (nrow(TPRf) + nrow(FNRf))
cat("Gaussian RBF Kernel Recall:" , recallRf, "\n")

```